A boundary-integral framework to simulate \\ the viscous erosion of a porous medium

### 2018-04-05 docs/methods/erosion_methods.tex
NM

Flow-induced erosion reshapes solid material over a range of scales found in nature, from large-scale land formations \cite{han1969, Rothman2012, coh-dev-sey-yi-szy-rot2015} down to human-scale features and patterns \cite{daerr2003erosion}. Though less visible, these same forces are at work at the very smallest scales, slowly deteriorating the individual constituents of porous media (e.g.~solid, sand, or clay) or biological structures like plaque and biofilms \cite{pic-van-hei2000, sha2002, gro-gij-van-fer-hat-van-yua-wen2007}. Motivated by such small-scale applications, we develop here numerical methods to simulate fluid-mechanical erosion of multiple solid bodies in Stokes flow --- the most relevant regime for groundwater and bio-fluid applications.

Recent work in the high-Reynolds regime has highlighted the importance of the feedback between flow and morphology that occurs during erosion \cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, hewett2017evolution, moore2017riemann}, as well as during related processes such as dissolution and melting \cite{Huang2015, kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion, hewett2017pear, claudin2017dissolution}. In such cases, the flow alters the shape of the immersed structures, which in turn modifies the surrounding flow, and so on. Collectively, such processes have been referred to as `sculpting with flow'~\cite{ristroph2018sculpting}. From this perspective, the example of an eroding porous medium presents a new opportunities and challenges. First, the medium itself is composed of many solid constituents, and so multi-body interactions, especially close interactions, are key. Second, the long-range interactions facilitated by Stokes flow even more strongly couples the shape evolution of different bodies.

The methods developed here will allow us to simulate the erosion of many bodies, as demonstrated in Fig.~\ref{fig:50bodies} in the case of 50 bodies of randomly assigned sizes and positions. In this work, mechanical erosion arises from the shear-stress induced by the flow, as has been experimentally confirmed \cite{ris-moo-chi-she-zha2012}. In Fig.~\ref{fig:50bodies}, color represents the vorticity of the surrounding flow. When evaluated on the solid boundaries, vorticity is equivalent to shear, and thus provides a convenient tool for visualizing local erosion rates. As shown in the figure, our simulations allows solid bodies to diminish in size and vanish in finite time, while other larger bodies persist.

>>>>

Flow-induced erosion deteriorates and reshapes solid material over a range of scales found in nature, from large-scale land formations \cite{han1969, Rothman2012, coh-dev-sey-yi-szy-rot2015} down to human-scale features and patterns \cite{daerr2003erosion}. Though less visible, these same forces are at work at the very smallest scales, slowly deteriorating the individual constituents of porous media (e.g.~solid, sand, or clay) or biological structures like plaque and biofilms \cite{pic-van-hei2000, sha2002, gro-gij-van-fer-hat-van-yua-wen2007}. Motivated by such small-scale applications, we develop here numerical methods to simulate fluid-mechanical erosion of multiple solid bodies in Stokes flow --- the most relevant regime for groundwater and bio-fluid applications.

Recent work in the high-Reynolds regime has highlighted the importance of the feedback between flow and morphology that occurs during erosion \cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, hewett2017evolution, moore2017riemann}. A similar feedback mechanism is at work in the related processes of dissolution and melting \cite{Huang2015, kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion, hewett2017pear, claudin2017dissolution}. In such cases, the flow alters the shape of the immersed structures, which in turn modifies the surrounding flow, and so on. Collectively, such processes have been referred to as `sculpting with flow'~\cite{ristroph2018sculpting}. From this viewpoint, the example of an eroding porous medium presents a new opportunities and challenges. First, the medium itself is composed of many solid constituents, and so multi-body interactions, especially close interactions, must be considered. Second, the long-range interactions facilitated by Stokes flow even more strongly couples the shape evolution of multiple bodies.

The methods developed here will allow us to simulate the erosion of many bodies, as demonstrated in Fig.~\ref{fig:50bodies} in for a medium composed of 50 bodies with randomly assigned initial positions and sizes. In this work, mechanical erosion arises from the shear-stress induced by the flow, as has been experimentally confirmed \cite{ris-moo-chi-she-zha2012}. In Fig.~\ref{fig:50bodies}, color represents the vorticity of the surrounding flow. When evaluated on the solid boundaries, vorticity is equivalent to shear, and thus provides a convenient tool for visualizing local erosion rates. As shown in the figure, our simulations allows solid bodies to diminish in size and vanish in finite time, while other larger bodies persist.
